Software Architectures for Integrating Workflow Engines in Science Gateways

### 2016-05-13 paper.tex
done 3.4

\paragraph{Robustness} No component is specific to workflow execution
(\texttt{R$_1$}=0), and only interaction \texttt{e} is
(\texttt{R$_2$}=1)\closedanswerednote{sorina}{I understand why this is 1
  (and why interactions c and b are not counted here), but I wonder
  whether we can simply ignore these interactions because they are not
  specific to workflow execution. As far as robustness is concerned,
  the transfers have to be successful anyway...}{Tristan}{We cannot
  just ignore these interactions in the architecture because they are
  required for the system to work. The transfers have to be successful
  but data movements in the workflow will use the same code as data
  movements done in the science gateway, which will make this
  architecture more robust than service for instance. In VIP, for
  instance, the code used in the GASW wrapper to transfer data (c2) is
  completely independent from the ``file transfer'' module in the
  platform (c1), which is detrimental to robustness}{Tristan}{Closed,
  see github issue \#7.}.

\paragraph{Extensibility} Integrating a new type of workflow engine
requires to develop interaction \texttt{e} and to install the engine
on the infrastructure (\texttt{E$_1$}=2). Updating an engine version
in the architecture shares the same mechanism as version updates of
other tasks, which requires an update on the
infrastructure (\texttt{E$_2$}=1).  New workflows are integrated by
creating a new task in the science gateway through interactions
\texttt{a$_1$} and \texttt{a$_2$} (\texttt{E$_3$}=2). Adding a new
infrastructure requires to update interactions \texttt{a$_1$},
\texttt{b} and \texttt{c} in the science gateway (\texttt{E$_4$}=3).

@@

\paragraph{Scalability}
New engine instances are spawned and executed on the infrastructure as
any other task upon user submission (\texttt{S$_1$}=0). This is a
major interest of task encapsulation. Distributed engines are not
supported by default (\texttt{S$_2$}=1). Task scheduling is slightly
more complex than in the other approaches due to the special role of
the task that executes the workflow engine (\texttt{S$_3$}=1). Indeed,
the reliability of this task is critical since all the sub-tasks in
the workflow depend on it and, depending on the recovery capabilities
of the workflow engine, may need to be resubmitted if the workflow
task fails. The workflow task is also longer than all its sub-tasks,
which increases its chances of failure. In addition, task parameters,
for instance estimated walltime, are more difficult to estimate for
the workflow task than for the sub-tasks because workflows are by
definition more complex than their sub-tasks: errors on sub-task
parameter estimations accumulate in the workflow, and additional
control constructs such as tests and loops may further increase the
uncertainty. Such parameter estimation errors may generate issues such
as selection of wrong batch queues on clusters or task termination due
to exceeded quotas. \closednote{Naj}{The lay reader wonders why?  an
  example like the one offered for interdependencies can
  help}{Tristan}{Done, thanks.} Finally, the interdependencies between
the workflow task and its sub-tasks may create deadlocks when there is
contention. For instance, if only 1 computing resource is available
for the science gateway and if the workflow task is running on it and
submits sub-tasks, then the sub-tasks could only execute when the
resource is available, which will never happen because the workflow
task will not complete until the sub-tasks complete. This
configuration can be generalized to an infrastructure with $n$
resources where $n$ workflows are submitted. In practice, however, the
number of submitted workflows usually remains lower than the number of
computing resources available on this infrastructure, which makes such
deadlocks unlikely to happen. \closednote{Naj}{so is this the solution to
  fixing the above mentioned limitations with
  scalability?}{Tristan}{Unfortunately this is not a solution. We just
  say that this situation is unlikely to happen which is true.}

\paragraph{Specific features} Meta-workflows are not available
(\texttt{O$_1$}=1).  Obtaining fine-grained information about workflow
tasks is not straightforward since the science gateway has no
knowledge about the workflow topology, and the workflow engine is
integrated as a task (\texttt{O$_2$}=1).

>>>>

\paragraph{Robustness} No component is specific to workflow execution
(\texttt{R$_1$}=0), and only interaction \texttt{e} is necessary
(\texttt{R$_2$}=1)\closedanswerednote{sorina}{I understand why this is 1
  (and why interactions c and b are not counted here), but I wonder
  whether we can simply ignore these interactions because they are not
  specific to workflow execution. As far as robustness is concerned,
  the transfers have to be successful anyway...}{Tristan}{We cannot
  just ignore these interactions in the architecture because they are
  required for the system to work. The transfers have to be successful
  but data movements in the workflow will use the same code as data
  movements done in the science gateway, which will make this
  architecture more robust than service for instance. In VIP, for
  instance, the code used in the GASW wrapper to transfer data (c2) is
  completely independent from the ``file transfer'' module in the
  platform (c1), which is detrimental to robustness}{Tristan}{Closed,
  see github issue \#7.}.

\paragraph{Extensibility} Integrating a new type of workflow engine
requires developing interaction \texttt{e} and  installing the engine
on the infrastructure (\texttt{E$_1$}=2). Updating an engine version
in the architecture shares the same mechanism as version updates of
other tasks, which requires an update on the
infrastructure (\texttt{E$_2$}=1).  New workflows are integrated by
creating a new task in the science gateway through interactions
\texttt{a$_1$} and \texttt{a$_2$} (\texttt{E$_3$}=2). Adding a new
infrastructure requires updating interactions \texttt{a$_1$},
\texttt{b} and \texttt{c} in the science gateway (\texttt{E$_4$}=3).

@@

\paragraph{Scalability}
New engine instances are spawned and executed on the infrastructure as
any other task upon user submission (\texttt{S$_1$}=0). This is a
major interest of task encapsulation. Distributed engines are not
supported by default (\texttt{S$_2$}=1). Task scheduling is slightly
more complex than in the other approaches due to the special role of
the task that executes the workflow engine (\texttt{S$_3$}=1). 
\note{sil}{i have the impression that the comments below are about robustness, and not so much about scalability. i also think that the level of detail here is much larger than in other 3.X sections. finally, i wonder whether what is explained here (which i must admit i could not follow completely in the scope of this evaluation or architectures) is valid only for this architecture in particular. i have the impression that this argument is more general }
Indeed,
the reliability of this task is critical since all the sub-tasks in
the workflow depend on it and, depending on the recovery capabilities
of the workflow engine, may need to be resubmitted if the workflow
task fails. The workflow task is also longer than all its sub-tasks,
which increases its chances of failure. In addition, task parameters,
for instance estimated walltime, are more difficult to estimate for
the workflow task than for the sub-tasks because workflows are by
definition more complex than their sub-tasks: errors on sub-task
parameter estimations accumulate in the workflow, and additional
control constructs such as tests and loops may further increase the
uncertainty. Such parameter estimation errors may generate issues such
as selection of wrong batch queues on clusters or task termination due
to exceeded quotas. \closednote{Naj}{The lay reader wonders why?  an
  example like the one offered for interdependencies can
  help}{Tristan}{Done, thanks.} Finally, the interdependencies between
the workflow task and its sub-tasks may create deadlocks when there is
contention. For instance, if only 1 computing resource is available
for the science gateway and if the workflow task is running on it and
submits sub-tasks, then the sub-tasks could only execute when the
resource is available, which will never happen because the workflow
task will not complete until the sub-tasks complete. This
configuration can be generalized to an infrastructure with $n$
resources where $n$ workflows are submitted. In practice, however, the
number of submitted workflows usually remains lower than the number of
computing resources available on this infrastructure, which makes such
deadlocks unlikely to happen. \closednote{Naj}{so is this the solution to
  fixing the above mentioned limitations with
  scalability?}{Tristan}{Unfortunately this is not a solution. We just
  say that this situation is unlikely to happen which is true.}

\paragraph{Specific features} Meta-workflows are not available
(\texttt{O$_1$}=1).  
\note{sil}{see comment in section 3.1}
Obtaining fine-grained information about workflow
tasks is not straightforward since the science gateway has no
knowledge about the workflow topology, and the workflow engine is
integrated as a task (\texttt{O$_2$}=1).
\note{sil}{mm, i think the case here is the same as for service. or actually i would give this one a better score, because there are less components to cross}
